A multi-dimensional extension of the Lightweight Temporal Compression method

### 2018-10-17 paper.tex
Added conclusion

% Compression of stream data
% need references
Compressing data streams is required in many contexts, to reduce 
storage and network requirements of data analytics. In particular, it is 
highly beneficial to energy-constrained devices such as Systems on 
Modules (SoM), as they reduce the amount of energy-consuming radio transmissions.
However, resource-intensive algorithms such as~\cite{x},~\cite{y} 
or~\cite{z} are not well-suited for energy-constrained systems, due to 
the limited memory available on these systems (typically a few KB), and 
the energy consumption associated with CPU usage. On such systems, 
compression algorithms need to find a trade-off between reducing 
network communications and increasing memory and CPU usage. As 
discussed in~\cite{zordan2014performance}, linear compression methods 
provide a very good compromise between these two factors, leading to 
substantial energy reduction.

@@

In this paper, we extend LTC to dimension $n$. To do so, we propose an 
algebraic formulation of the algorithm that also yields a 
norm-independent expression of it. We implement our extension on 
Motsai's Neblina module, and we test it on 3D acceleration streams 
acquired during human exercises, namely biceps curling, walking and 
running. Our implementation of LTC is available as free software on GitHub\todo{add link}.

We assume that the stream consists of a sequence of data points 
received at uneven intervals. The compression algorithm 
\emph{transmits} fewer points than it receives. The transmitted points 
might be part of the stream, or computed from stream elements. The 
\emph{compression ratio} is the ratio between the number of received 
points and the number of transmitted points. An application would 
reconstruct the stream from the transmitted points: the 
\emph{reconstruction error} is the maximum absolute difference between 
a data point of the reconstructed stream, and the corresponding data 
point in the original stream.

% Outline
Section~\ref{sec:ltc} provides some background on the LTC algorithm, 
and formalizes the description initially proposed 
in~\cite{schoellhammer2004lightweight}. Section~\ref{sec:extension} 
presents our norm-independent extension to dimension $n$, and 
Section~\ref{sec:implementation} describes our implementation. 
Section~\ref{sec:results} reports on experiments to validate our 
implementation, and evaluate the impact of n-dimensional LTC on SoM 
energy consumption.
\begin{figure}
\includegraphics[width=\columnwidth]{ltc.pdf}
\caption{Illustration of the LTC algorithm and notations used. Red dots indicate transmitted points. Dashed lines
represent the high line and the low line at times when a point is transmitted.}
\label{fig:ltc}
\end{figure}

@@

Euclidean norm would benefit from a better algorithm but it still
provides interesting energy savings.

\section*{Acknowledgement}
This work was funded by the Canadian

>>>>

% Compression of stream data
% need references
Lossy compression of data streams is required in many contexts, to reduce 
storage and network requirements of data analytics. In particular, it 
is highly beneficial to energy-constrained devices such as Systems on 
Modules (SoM), as they reduce the amount of energy-consuming radio 
transmissions. However, resource-intensive algorithms such as the ones 
based on polynomial interpolation or auto-regression 
methods~\cite{lu2010optimized} are not well-suited for 
energy-constrained systems, due to the limited memory available on 
these systems (typically a few KB), and the energy consumption 
associated with CPU usage. On such systems, compression algorithms need 
to find a trade-off between reducing network communications and 
increasing memory and CPU usage. As 
discussed in~\cite{zordan2014performance}, linear compression methods 
provide a very good compromise between these two factors, leading to 
substantial energy reduction.

@@

In this paper, we extend LTC to dimension $n$. To do so, we propose an 
algebraic formulation of the algorithm that also yields a 
norm-independent expression of it. We implement our extension on 
Motsai's Neblina module, and we test it on 3D acceleration streams 
acquired during human exercises, namely biceps curling, walking and 
running. Our implementation of LTC is available as free software.

We assume that the stream consists of a sequence of data points 
received at uneven intervals. The compression algorithm 
\emph{transmits} fewer points than it receives. The transmitted points 
might be included in the stream, or computed from stream points. The 
\emph{compression ratio} is the ratio between the number of received 
points and the number of transmitted points. An application would 
reconstruct the stream from the transmitted points: the 
\emph{reconstruction error} is the maximum absolute difference between 
a point of the reconstructed stream, and the corresponding 
point in the original stream.

\begin{figure}
\includegraphics[width=\columnwidth]{ltc.pdf}
\caption{Illustration of the LTC algorithm and notations used. Blue 
dots are received points, red dots are transmitted points. Dashed lines 
represent the high line and the low line at times when a point is 
transmitted.\vspace*{-0.3cm}}
\label{fig:ltc}
\end{figure}

% Outline
Section~\ref{sec:ltc} provides some background on the LTC algorithm, 
and formalizes the description initially proposed 
in~\cite{schoellhammer2004lightweight}. Section~\ref{sec:extension} 
presents our norm-independent extension to dimension $n$, and 
Section~\ref{sec:implementation} describes our implementation. 
Section~\ref{sec:results} reports on experiments to validate our 
implementation, and evaluate the impact of n-dimensional LTC on SoM 
energy consumption.

@@

We presented an extension of the Lightweight Temporal Compression 
method to dimension $n$ that can be instantiated for any distance 
function. Our extension formulates LTC as an intersection dection 
problem between n-balls. We implemented our extension on a SoM for the 
infinity and Euclidean norms, and we measured the energy reduction 
resulting from compression for acceleration streams acquired during a 
few human activities.

We conclude from our experiments that the proposed extension to LTC is 
well suited to reduce energy consumption on SoMs, as it reduce energy 
consumption by up to 20\%. The implementation behaves better with the 
infinity norm than with the Euclidean one, due to the lack of efficient 
algorithms to detect the intersection between n-balls defined for the 
Euclidean norms.

Our future work focuses on this issue. Our current implementation with 
the Euclidean norm relies on Helly's theorem, which only provides an 
algorithm of complexity ${m \choose n+1}$ to compress $m$ points in 
dimension $n$. We note that Helly's theorem works for arbitrary convex 
subsets of $\mathbb{R}^n$, while we are considering a \emph{sequence} 
of \emph{balls} of \emph{decreasing radius}. Based on this 
observation, a stronger result might exist that would lead to a more 
efficient implementation of LTC with the Euclidean norm. Our current 
idea is to look for a point function of the ball centers that would 
necessarily belong to the ball intersection when it is not empty; such 
a point, if it exists, necessarily converges to the center of the last 
ball in the sequence as $n$ increases, as the radius of the last ball
decreases to zero.

The effect of norm selection should not be underestimated. Some 
situations might be better described with the Euclidean norm than with the
infinity norm, such as the ones involving position or movement measures. 
Using the infinity norm instead of the Euclidean would lead to important 
error differences, proportional to $\sqrt{n}$ in dimension $n$.

\section*{Acknowledgement}
This work was funded by a Strategic Project Grant of the Natural 
Sciences and Engineering Research Council of Canada. We thank Hosein 
Nourani for providing the data used in Experiment 1.
